---
title: "ESM 206 Assignment 4 Problem 2"
author: "Robert Heim"
date: "11/14/2018"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include= FALSE}

 # Loaded tidyverse and other packages 

library(tidyverse)
library(pwr)
library(knitr)
library(kableExtra)
library(plotly)
library(extrafont)
library(ggrepel)
library(effsize)
library(vcdExtra)


```

# Problem 2: Compare Mean Lobster Size By Site in 2017

```{r include=FALSE}

# Read in dataset for lobster size 

lobster_size <- read_csv("lobster_size_abundance.csv")

lobster_size

```

```{r}

# Get data into case and tidy format

count <- as.data.frame(lobster_size) #Coerce to data.frame 

lobster_size_tidy <- expand.dft(count, freq = "COUNT") #Expand data to tidy format

lobster_size_tidy_2017 <- lobster_size_tidy %>% 
  filter(YEAR == 2017) #Use only observations only from 2017

```

```{r}

# Create a summary table for mean lobster carapace length in 2017 at each site

size_mean_2017 <- lobster_size_tidy_2017 %>%
  group_by(SITE) %>%
  summarize(Mean_size= mean(SIZE),Standard_deviation= sd(SIZE), Sample_size= length(SITE), Standard_error= Standard_deviation/sqrt(Sample_size))
  
```

```{r}

# Single factor: Site 
# Number of levels in the factor: 5 (AQUE, CARP, IVEE, MOHK, and NAPL)
# Random variable: carapace length (mm)

#Question: Is there a significant difference in carapace length (mm) for AQUE, CARP, IVEE, MOHK, and NAPL locations in 2017?

# H0: Mean lobster carapace length across all sites are equal. 
# HA: At least two mean lobster caparace lengths differ significantly between sites. 

# Do some data exploring to find whether the assumption of normality is met for the carapace length data. Create a histogram and qq-plot.

lobster_2017_hist <- lobster_size_tidy_2017 %>%
  ggplot(aes(x= SIZE)) +
  geom_histogram(binwidth = 1, aes(fill= SITE))+
  facet_wrap((~SITE)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,9))+ 
  scale_x_continuous(expand = c(0,0))

lobster_2017_hist

lobster_2017_qq <- lobster_size_tidy_2017 %>%
  ggplot(aes(sample= SIZE)) +
  geom_qq() +
  facet_wrap(~ SITE)

lobster_2017_qq

# Lobster size data for 2017 appear normally distributed at each site

# Use a one-way ANOVA to test for a difference between any two of the sites

lobster_aov <- aov(SIZE ~ SITE, data = lobster_size_tidy_2017)

summary(lobster_aov)

# P-value= 0.009, meaning at least two samples of lobster carapance length were taken from sites with different means. Which ones are different? Are all five different from each other? Or something else? 

# Need to run a Post-hoc Tukey's test to find where the differences are. 

lobster_ph <- TukeyHSD(lobster_aov) 

lobster_ph


```

Mean lobster carapace length in 2017 differed significantly at the five Long-Term Ecological Resesearch Sites (one-way ANOVA, F(4,1663)= 3.42, p= 0.009; Figure 1). Post hoc analysis by Tukey's HSD revealed that carapace length differed significantly between Naples Reef and Carpinteria (pairwise p= 0.023). Carapace length also differed significantly between Naples Reef and Isla Vista (0.004). 

“Sediment bulk densities (g/cm3
) differed significantly in the four river reaches studied (one-way ANOVA,
F(3,92) = 85.0, p < 0.001, n = 24 for each; Table 1). Post-hoc analysis by Tukey’s HSD revealed that
sediment density in Reach A differed significantly from all other reaches (pairwise p < 0.001) and had the
lowest mean bulk density (0.38 g/cm3
). Sediment bulk densities in the downstream reaches were more than
double that measured in Reach A (Figure 1), indicating that Reach A may currently offer the most suitable
spawning habitat.”

```{r}

# Use a bar graph to show mean lobster carapace length at each site. Include error bars and letters.  

	
lobster_size_col <-ggplot(size_mean_2017, aes(x = SITE, y = Mean_size)) +
  geom_col(aes(fill = SITE)) +
  geom_errorbar(aes(ymin= Mean_size-Standard_error, ymax= Mean_size+Standard_error), width=.2) + 
  theme_classic() +
  theme(legend.position= "none") +
  labs(x= "Location", y= expression(Mean~Carapace~Length~(mm))) +
  scale_y_continuous(limit= c(0,100), expand=c(0,0)) +
  scale_x_discrete(labels=c("AQUE" = "Arroyo Quemado\nn=67", 
                            "NAPL" = "Naples Reef\nn=705",
                            "MOHK" = "Mohawk Reef\nn=606",
                            "IVEE" = "Isla Vista\nn=178",
                            "CARP" = "Carpinteria\nn=112")) +
  annotate("text", x = 1, y = 87, label = "A, B", family = "Times New Roman") +
  annotate("text", x = 2, y = 87, label = "A", family = "Times New Roman") +
  annotate("text", x = 3, y = 87, label = "A", family = "Times New Roman") +
  annotate("text", x = 4, y = 87, label = "A, B", family = "Times New Roman") +
  annotate("text", x = 5, y = 87, label = "B", family = "Times New Roman")

lobster_size_col



```

**Figure 1. Mean Lobster Carapace Length $(mm)$ at five sites in the Santa Barbara Channel in 2017.** Error bars represent ± 1 standard error. Like letters indicate values that do not differ significantly. Data retreived from Santa Barbara Coastal Long-Term Ecological Research [1]. 


# 3 Changes in lobster size at each MPA and non-MPA site (comparing only 2012 and 2017 sizes). 

```{r}

# Use a one-sided, two-sample t-test to compare lobster size between 2012-2017 at each MPA site

#HO: Lobsters at IVEE are not significantly larger in 2017 than in 2012
#HA: Lobsters at IVEE are significantly larger in 2017 than in 2012

# Create a new df with lobster size for IVEE at 2012 and 2017 

lobster_size_IVEE_2012 <- lobster_size_tidy %>% 
  filter(SITE == "IVEE") %>% 
  filter(YEAR == "2012")  
 
lobster_size_IVEE_2017 <- lobster_size_tidy %>% 
  filter(SITE == "IVEE") %>% 
  filter(YEAR == "2017") 

# Find the mean and sd for IVEE for 2012 and 2017

IVEE_2017_mean <- mean(lobster_size_IVEE_2017$SIZE)

IVEE_2017_sd <-sd(lobster_size_IVEE_2017$SIZE)
 
IVEE_2012_mean <- mean(lobster_size_IVEE_2012$SIZE)

IVEE_2012_sd <-sd(lobster_size_IVEE_2012$SIZE)

IVEE_2017_mean - IVEE_2012_mean
 
# Run a one-sided t-test for each MPA, comparing 2012 to 2017 

lobster_IVEE_ttest<- t.test(lobster_size_IVEE_2017$SIZE, lobster_size_IVEE_2012$SIZE, var.equal= TRUE, alternative= "greater")

lobster_IVEE_ttest

# Need to do an effect size test using Cohen's D
# First, make vectors

IVEE_2017_vector <- lobster_size_IVEE_2017 %>%
  pull(SIZE)

IVEE_2012_vector <- lobster_size_IVEE_2012 %>%
  pull(SIZE)

effect_size <- cohen.d(IVEE_2017_vector,IVEE_2012_vector)
effect_size



```
Mean lobsters carapace length in 2017  [`r round(IVEE_2017_mean,2)` ±  `r round(IVEE_2017_sd,2)` $mm$ ($\mu$ ± $\sigma$), n=26] was not signficantly different than in 2012 [`r round(IVEE_2012_mean,2)` ±  `r round(IVEE_2012_sd,2)` $mm$ ($\mu$ ± $\sigma$), n=606] at Isla Vista [t(`r round(lobster_IVEE_ttest$parameter,2)`) = `r round(lobster_IVEE_ttest$statistic,2)`, p < `r round(lobster_IVEE_ttest$p.value,3)`, $\alpha$ = 0.05]. The effect size is moderate (Cohen's d= 0.38). From the samples, lobster carapace length in 2017 was approximately 5.38 mm greater than lobster caparace length in 2012. 


```{r}

# Run a one-sided, two-sample t-test to see if size in 2017 is greater than in 2012

lobster_size_NAPL_2012 <- lobster_size_tidy %>% 
  filter(SITE == "NAPL") %>% 
  filter(YEAR == "2012")  
 
lobster_size_NAPL_2017 <- lobster_size_tidy %>% 
  filter(SITE == "NAPL") %>% 
  filter(YEAR == "2017") 

# Find the mean and sd for NAPL for 2012 and 2017

NAPL_2017_mean <- mean(lobster_size_NAPL_2017$SIZE)

NAPL_2017_sd <-sd(lobster_size_NAPL_2017$SIZE)
 
NAPL_2012_mean <- mean(lobster_size_NAPL_2012$SIZE)

NAPL_2012_sd <-sd(lobster_size_NAPL_2012$SIZE)

NAPL_2017_mean - NAPL_2012_mean

# Run a one-sided t-test for to see if size in 2017 is greater than in 2012

lobster_NAPL_ttest<- t.test(lobster_size_NAPL_2017$SIZE, lobster_size_NAPL_2012$SIZE, var.equal= TRUE)

lobster_NAPL_ttest

# Lobster sizes at NAPL in 2017 are not significantly different than in 2012

NAPL_2017_vector <- lobster_size_NAPL_2017 %>%
  pull(SIZE)

NAPL_2012_vector <- lobster_size_NAPL_2012 %>%
  pull(SIZE)

effect_size <- cohen.d(NAPL_2017_vector,NAPL_2012_vector)
effect_size

```
Mean lobsters carapace length in 2017  [`r round(NAPL_2017_mean,2)` ±  `r round(NAPL_2017_sd,2)` $mm$ ($\mu$ ± $\sigma$), n=6] was not signficantly different than in 2012 [`r round(NAPL_2012_mean,2)` ±  `r round(NAPL_2012_sd,2)` $mm$ ($\mu$ ± $\sigma$), n=112] at Naples Reef [t(`r round(lobster_IVEE_ttest$parameter,2)`) = `r round(lobster_NAPL_ttest$statistic,2)`, p < `r round(lobster_NAPL_ttest$p.value,3)`, $\alpha$ = 0.05]. The effect size is small (Cohen's d= 0.28). From the samples, lobster carapace length in 2017 was approximately 3.23 mm greater than lobster caparace length in 2012. 

```{r}

# One-sided t-test for AQUE nonMPA 

lobster_size_AQUE_2012 <- lobster_size_tidy %>% 
  filter(SITE == "AQUE") %>% 
  filter(YEAR == "2012")  

lobster_size_AQUE_2017 <- lobster_size_tidy %>% 
  filter(SITE == "AQUE") %>% 
  filter(YEAR == "2017") 

lobster_AQUE_ttest<- t.test(lobster_size_AQUE_2017$SIZE, lobster_size_AQUE_2012$SIZE, var.equal= TRUE)

lobster_AQUE_ttest

# Find the mean and sd for AQUE for 2012 and 2017

AQUE_2017_mean <- mean(lobster_size_AQUE_2017$SIZE)

AQUE_2017_sd <-sd(lobster_size_AQUE_2017$SIZE)
 
AQUE_2012_mean <- mean(lobster_size_AQUE_2012$SIZE)

AQUE_2012_sd <-sd(lobster_size_AQUE_2012$SIZE)

AQUE_2017_mean - AQUE_2012_mean

# Lobster sizes at AQUE in 2017 are not significantly different than in 2012

AQUE_2017_vector <- lobster_size_AQUE_2017 %>%
  pull(SIZE)

AQUE_2012_vector <- lobster_size_AQUE_2012 %>%
  pull(SIZE)

effect_size <- cohen.d(AQUE_2017_vector,AQUE_2012_vector)
effect_size


```
Mean lobsters carapace length in 2017  [`r round(AQUE_2017_mean,2)` ±  `r round(AQUE_2017_sd,2)` $mm$ ($\mu$ ± $\sigma$), n=38] was not signficantly different than in 2012 [`r round(AQUE_2012_mean,2)` ±  `r round(AQUE_2012_sd,2)` $mm$ ($\mu$ ± $\sigma$), n=67] at Arroyo Quemado [t(`r round(lobster_AQUE_ttest$parameter,2)`) = `r round(lobster_AQUE_ttest$statistic,2)`, p < `r round(lobster_AQUE_ttest$p.value,3)`, $\alpha$ = 0.05]. The effect size is small (Cohen's d= 0.26). From the samples, lobster carapace length in 2017 was approximately 2.90 mm greater than lobster caparace length in 2012. 

```{r}

# One-sided t-test for CARP nonMPA 

lobster_size_CARP_2012 <- lobster_size_tidy %>% 
  filter(SITE == "CARP") %>% 
  filter(YEAR == "2012")  

lobster_size_CARP_2017 <- lobster_size_tidy %>% 
  filter(SITE == "CARP") %>% 
  filter(YEAR == "2017") 

lobster_CARP_ttest<- t.test(lobster_size_CARP_2017$SIZE, lobster_size_CARP_2012$SIZE,var.equal= TRUE)

lobster_CARP_ttest

# Find the mean and sd for CARP for 2012 and 2017

CARP_2017_mean <- mean(lobster_size_CARP_2017$SIZE)

CARP_2017_sd <-sd(lobster_size_CARP_2017$SIZE)
 
CARP_2012_mean <- mean(lobster_size_CARP_2012$SIZE)

CARP_2012_sd <-sd(lobster_size_CARP_2012$SIZE)

CARP_2017_mean - CARP_2012_mean

# Lobster sizes at CARP in 2017 are not significantly different than in 2012

CARP_2017_vector <- lobster_size_CARP_2017 %>%
  pull(SIZE)

CARP_2012_vector <- lobster_size_CARP_2012 %>%
  pull(SIZE)

effect_size <- cohen.d(CARP_2017_vector,CARP_2012_vector)
effect_size


# Lobster sizes at CARP in 2017 are not significantly greater than in 2012
```
Mean lobsters carapace length in 2017  [`r round(CARP_2017_mean,2)` ±  `r round(CARP_2017_sd,2)` $mm$ ($\mu$ ± $\sigma$), n=78] was not signficantly different than in 2012 [`r round(CARP_2012_mean,2)` ±  `r round(CARP_2012_sd,2)` $mm$ ($\mu$ ± $\sigma$), n=705] at Carpinteria [t(`r round(lobster_AQUE_ttest$parameter,2)`) = `r round(lobster_CARP_ttest$statistic,2)`, p < `r round(lobster_CARP_ttest$p.value,3)`, $\alpha$ = 0.05]. The effect size is small (Cohen's d= -0.16). From the samples, lobster carapace length in 2017 was approximately 2.13 mm less than lobster caparace length in 2012. 

```{r}
# One-sided t-test for MOHK nonMPA 

lobster_size_MOHK_2012 <- lobster_size_tidy %>% 
  filter(SITE == "MOHK") %>% 
  filter(YEAR == "2012")  

lobster_size_MOHK_2017 <- lobster_size_tidy %>% 
  filter(SITE == "MOHK") %>% 
  filter(YEAR == "2017") 

lobster_MOHK_ttest<- t.test(lobster_size_MOHK_2017$SIZE, lobster_size_MOHK_2012$SIZE,var.equal= TRUE, alternative = "less")

lobster_MOHK_ttest

# Lobster sizes are signficantly different at MOHK in 2012 and 2017

# Find the mean and sd for MOHK for 2012 and 2017

MOHK_2017_mean <- mean(lobster_size_MOHK_2017$SIZE)

MOHK_2017_sd <-sd(lobster_size_MOHK_2017$SIZE)
 
MOHK_2012_mean <- mean(lobster_size_MOHK_2012$SIZE)

MOHK_2012_sd <-sd(lobster_size_MOHK_2012$SIZE)

MOHK_2017_mean - MOHK_2012_mean

# Lobster sizes at MOHK in 2017 are not significantly different than in 2012

MOHK_2017_vector <- lobster_size_MOHK_2017 %>%
  pull(SIZE)

MOHK_2012_vector <- lobster_size_MOHK_2012 %>%
  pull(SIZE)

effect_size <- cohen.d(MOHK_2017_vector,MOHK_2012_vector)
effect_size

```
Mean lobsters carapace length in 2017  [`r round(MOHK_2017_mean,2)` ±  `r round(MOHK_2017_sd,2)` $mm$ ($\mu$ ± $\sigma$), n=83] was signficantly less than in 2012 [`r round(MOHK_2012_mean,2)` ±  `r round(MOHK_2012_sd,2)` $mm$ ($\mu$ ± $\sigma$), n=178] at Mohawk Reef [t(`r round(lobster_MOHK_ttest$parameter,2)`) = `r round(lobster_MOHK_ttest$statistic,2)`, p < `r round(lobster_MOHK_ttest$p.value,3)`, $\alpha$ = 0.05]. The effect size is moderate (Cohen's d= -0.54). From the samples, lobster carapace length in 2017 was approximately 5.25 mm less than lobster caparace length in 2012. 

#4 Proportion of legal lobsters at the five sites

```{r}

# Make a new data frame with numbers of lobster that are and are not above the legal minimum of 82.6 mm.  

legal_2017 <- lobster_size_tidy_2017 %>%
  mutate(above_legal= case_when(SIZE < 82.6 ~ 'No', SIZE >= 82.6 ~ 'Yes')) %>%
  count(SITE, above_legal) %>%
  spread(above_legal, n) %>%
  select(-SITE)

legal_2017

# Rename rows in legal_2017

rownames(legal_2017) <- c("Arroyo Quemado", " Carpinteria", "Isla Vista", "Mohawk Reef", "Naples Reef")

# Question:  Does the proportion of lobsters of legal size differ significantly across the five sites?

# Run a chi-square test

legal_2017_prop <- prop.table(as.matrix(legal_2017), 1)
size_x2 <- chisq.test(legal_2017)
size_x2

legal_2017_prop

# Find the proportion of lobsters at each site that are above the legal minimum of 82.6 mm.

size_prop <- prop.table(as.matrix(legal_2017_prop), 1)
size_prop 

```
Based on the samples from each Long-Term Ecological Research Site in 2017: Arroyo Quemado (n=67), Carpinteria (n=705), Isla Vista (n=606), Mohawk Reef (n=178), and Naples Reef (n=112), the proportion of lobsters that are above the legal minimum carapace length of 82.6 mm differs significantly between sites (X-squared(4) = 18.497, p< 0.001). 

